How fluid-mechanical erosion creates anisotropic porous media

### 2022-06-16 docs/PhysicsPaper/StokesErosion.tex
NM

The effects of flow-induced erosion are visible across a range of scales in the natural world, from massive geological structures \cite{abrams2009growth, perkins2015amplification, mac2020ultra, sharma2022alcove, mac2022morphological}, to mesoscopic patterns \cite{berhanu2012shape, bertagni2021hydrodynamic}, and down to the granular constituents comprising porous media \cite{zareei2022temporal}. Porous media encountered in natural settings typically exhibit {\em anisotropic} material properties,

\vsp{5}
% ORIGINAL BELOW
Effects of flow-induced erosion are visible across a range of scales, from massive geological structures, to mesoscopic patterns, and down to the tiny granular constituents that comprise a porous medium. 
In the case of a porous medium specifically, it has been long recognized that bulk properties of mediums encountered in nature are typically anisotropic, so that the medium allows flow in a preferred direction more easily. 
This anisotropy is commonly attributed to how grains are deposited, with their longest dimension parallel to the settling bed, so that they allow flow preferentially in the horizontal direction. Controlled experiments, however, have not been performed to test this hypothesis. Here, we use highly-accurate numerical simulations to examine an alternative, and possibly complementary, mechanism: namely, that the flow-induced erosion of the medium's solid constituents contributes to its overall anisotropy.

>>>>

Flow-induced erosion acts across a range of scales in the natural world, from massive geological structures sculpted by wind or water \cite{abrams2009growth, perkins2015amplification, mac2020ultra, sharma2022alcove, mac2022morphological}, to mesoscopic patterns formed by surface or internal flows \cite{berhanu2012shape, bertagni2021hydrodynamic}, and down to the granular constituents of porous media as they slowly disintegrate in groundwater flows \cite{zareei2022temporal}. The shape-flow feedback of erosion can imprint across all of these scales, affecting both large-scale features and small-scale ones, such as the microstructure of porous materials.
The porous media encountered in nature, for example, typically exhibit material {\em anisotropy} in that they permit seepage flow in certain directions more easily than in others. Typically, materials can be more permeable to flow in the longitudinal (or horizontal) direction than the transverse one by a factor of 5-20 \cite{bear1988dynamics, anderson2015applied}. Most commonly, this material anisotropy is attributed to the sedimentation process, in which, due to the fluid-structure interaction, non-spherical particles tend to settle with their long axis parallel to the plan of deposition \cite{bear1988dynamics}. Controlled experiments, however, have not been performed to test this hypothesis, and other mechanisms may also be at work. Here, we use highly-accurate numerical simulations to examine an alternative, and possibly complementary, mechanism: namely, that the flow-induced erosion of the medium's solid constituents contributes to its overall anisotropy.

\vsp{5}
